
---
author: "Eric Kernfeld"
output: html_document
---

###Setup 

This chunk sets up the file system and some functions to retrieve reference tables. It also loads in some useful packages and sets up a few hand-picked color schemes. 

You may need to edit the first few lines depending on where your data reside(s).

```{r}

PATH_TO_DATA = file.path( "~", "Dropbox", "2016JULY07scRNAseq", "data" )
# PATH_TO_METADATA = file.path( PATH_TO_DATA, "DROPseq_stats_working.csv")

# Generally indispensible
library(ggplot2)

UMAP_AXES = c("UMAP1", "UMAP2")


OVERVIEW_IDENT_INFO  = read.csv(file.path(proj_dir, "tables", "cluster_names.csv"), stringsAsFactors = F)
OVERVIEW_IDENT_INFO$name = OVERVIEW_IDENT_INFO$name_pth_grant_old
OVERVIEW_IDENT_COLORS =
  c(
    setNames(OVERVIEW_IDENT_INFO$colour, OVERVIEW_IDENT_INFO$ident %>% as.character), 
    setNames(OVERVIEW_IDENT_INFO$colour, OVERVIEW_IDENT_INFO$name), 
    setNames(OVERVIEW_IDENT_INFO$colour, OVERVIEW_IDENT_INFO$terminal_cluster)
  )

add_cluster_names = function(dge) {
  dge@meta.data$cluster_name     = setNames(OVERVIEW_IDENT_INFO$name,
                                            OVERVIEW_IDENT_INFO$ident)[as.character(dge@ident)]
  dge@meta.data$terminal_cluster = setNames(OVERVIEW_IDENT_INFO$terminal_cluster,
                                            OVERVIEW_IDENT_INFO$ident)[as.character(dge@ident)]
  dge
}


OVERVIEW_IDENT_ORDER = c(25, 4, 9, 26, 2, 12, 22, 17, 13, 
                         14, 16, 1, 21, 19, 8, 3,
                         10, 23, 24, 27, 15, 5, 0, 11,
                         6, 20, 7, 18) %>% as.character


fix_merged_atlas_labels = function(x) {
  x %<>% as.character
  old = c(25, 4, 26, 12, 22, 17, 24, 27, 21, 19, 0, 5, 6, 7, 18, 20)
  new = c("thymus__1", "thymus__2",
          "parathyroid", "ubb",
          "thyroid__1", "thyroid__2",
          "lung__1", "lung__2",
          "first_pouch__1", "first_pouch__2", 
          "trachea_esophagus" %>% paste0("__", 1:6))
  remaining_clusters = setdiff(0:27, old)
  old = c(old, remaining_clusters) %>% as.character
  new = c(new, paste0("PE__", seq_along(remaining_clusters))) 
  for( ii in seq_along(old)){
    x[x==old[[ii]]] = new[[ii]]
  }
  x
}


add_pouch_subsets = function(dge ){
  pouch_subsets = tryCatch(
    read.csv(inventory_get("pouch3_subsets"), row.names = 1), 
    error = function(e){ 
      stop(paste(
        "This function only works if 'pouch3_subsets'", 
        "has been added to the inventory.\n"
      ))
    } 
  )
  dge %<>% AddMetaData(pouch_subsets[c("Il7pos", "pouch3", "pouch34")])
}

#' This is an early version of thymusatlastools::get_batch_basis.
#' The two should give highly similar results, but to keep it
#' exactly backwards compatible with what I sent Hana, this 
#' version is maintained.
#'
get_batch_basis_hardwired = function( dge ) {
  # # Regress out a basis that contrasts between replicates but not across time points
  dge@meta.data$orig.ident %<>% droplevels()
  basis_all_reps = model.matrix( ~orig.ident + 0, dge@meta.data ) %>% as.data.frame
  
  batch_effect_indicators = with( basis_all_reps, data.frame(
    basis_e12  = orig.identCR_pharynx_E12_5_rep2 - orig.identCR_pharynx_E12_5_rep1,
    basis_e9   = orig.identCR_pharynx_E9_5_rep2  - orig.identCR_pharynx_E9_5_rep1,
    basis_e10a = orig.identCR_pharynx_E10_5_rep6 - orig.identCR_pharynx_E10_5_rep5,
    basis_e10b = orig.identCR_pharynx_E10_5_rep7 - orig.identCR_pharynx_E10_5_rep5,
    basis_e11a = orig.identCR_pharynx_E11_5_rep3 - orig.identCR_pharynx_E11_5_rep2B,
    basis_e11b = orig.identCR_pharynx_E11_5_rep2 - orig.identCR_pharynx_E11_5_rep2B
  )) %>% set_rownames(rownames(dge@meta.data))
  return( batch_effect_indicators )
}

# Retrieve Hana's analysis results.
#
get_hana_files = function(look_where = file.path(proj_dir, "related", "hana")){
  hana_file_names = list.files(look_where, full.names = T)
  hana_files = lapply(hana_file_names, read.delim, stringsAsFactors = F, sep = "" )
  names(hana_files) = basename(hana_file_names)
  for( ii in names(hana_files) ){
    if(grepl("foxn1|rank_genes", ii, ignore.case = T)){
      next
    }
    rownames(hana_files[[ii]]) = hana_files$obs$index
  }
  hana_files
}
```

```{r}
#' Plot one Seurat object (colorfully) on top of another (gray) using a shared coordinate system.
#'
knockout_plot = function(dge_background, 
                         dge_ko, 
                         colour, results_path = Sys.getenv("FREEZR_DESTINATION"), 
                         cols.use = c("khaki1", "red"), suffix = "",
                         do_projection = T,
                         do_hex_background = F, 
                         do_hex_ko = F,
                         do_density_ko = F,
                         ...,            
                         width = 8,
                         height = 7,
                         num_bins = 50){
  if(do_projection){
    dge_ko = ProjectCells(dge_train = dge_background, 
                          dge_test = dge_ko, 
                          regressout = "nUMI", 
                          to_project = UMAP_AXES)
  } 

  fields_used = c(UMAP_AXES, colour)
  hexify = function(df) bubblegrid::bubblegrid(df[[1]], df[[2]], df[[3]], num_bins = num_bins )$plot_env$hex_data
  pretend_to_hexify = function(df) {df$count = 1; df$x = df[[1]]; df$y = df[[2]]; df$color = df[[3]]; return(df)}
  background = FetchDataZeroPad(dge_background, fields_used )
  knockout   = FetchDataZeroPad(dge_ko,         fields_used )
  if(do_hex_background){
    background %<>% hexify
    background_mapping = aes(x = x, y = y, size = count)
  } else {
    background %<>% pretend_to_hexify
    background_mapping = aes(x = x, y = y)
  }
  if(do_hex_ko){
    knockout %<>% hexify
    knockout_mapping = aes(x = x, y = y, size = count, colour = color )
  } else {
    knockout %<>% pretend_to_hexify
    knockout_mapping = aes(x = x, y = y, colour = color )
  }
  if(is.numeric(knockout$color) & !is.factor(knockout$color)){
    make_colorscale = function(colours, ...) scale_color_gradientn( colours = colours, ... ) 
  } else {
    make_colorscale = function(colours, ...) scale_color_manual( values = colours, ... )
  }
  p = ggplot() + 
    geom_point(data = background, 
               mapping = background_mapping, colour = "gray" ) + 
    make_colorscale(colours = cols.use) + 
    coord_fixed() + theme(axis.text = element_blank(), 
                          axis.line = element_blank(), 
                          axis.ticks = element_blank() , 
                          axis.title = element_blank() ) 
  
  if( do_density_ko ){
    p = p + stat_density_2d(data = knockout[sample(nrow(knockout)), ],   
                            geom = "polygon", aes(x=x, y=y, alpha = ..level.., fill = color),
                            ...)

  } else {
    p = p + geom_point(data = knockout[sample(nrow(knockout)), ],   
                   mapping = knockout_mapping, ...) 
    
      
  }
  if(!is.null(results_path)){
    ggsave(file.path(results_path, paste0(colour, suffix, ".png")), width = width, height = height)
  }
  return(p)
}

superimpose_umap = knockout_plot

```


```{r}


get_hollander_foxn1_targets = function(x){
  return( read.csv(file.path(proj_dir, "tables", "Foxn1_high_confidence_genelist.txt"), stringsAsFactors = F )[[1]])
}
```

###Utilities

```{r}
rename_clusters_by_eday = function( dge, results_path, ident.use = "ident", new.name = "ident" ){
  days_present = unique(dge@meta.data$eday)
  get_eday_props = function(x) x %>% c(days_present) %>% table(exclude=NULL) %>% subtract(1) %>% prop.table
  make_eday_names = function( props ) props %>% is_greater_than(0.2) %>% which %>% names %>% paste0(collapse = "_")
  new_names = aggregate_nice( FetchData( dge, "eday"), 
                              by = FetchData(dge, ident.use), 
                              FUN = function(x) x %>% get_eday_props %>% make_eday_names ) 
  old_names = rownames(new_names)
  new_names %<>% c( new_names )
  add_numbers = function( x, sep = "_" ) {
    ave( seq_along(x), x, FUN = function(y) sort(order(y)) ) %>% paste0( x, sep, .  )
  }
  new_names %<>% add_numbers(sep = "_c")
  names(new_names) = old_names
  if( new.name == "ident"){
    dge %<>% SetIdent( new_names[dge@ident %>% as.character], cells.use = dge@cell.names )
  } else {
    dge %<>% AddMetaData( new_names[dge@ident %>% as.character] %>% setNames( dge@cell.names ), col.name = new.name )
  }
  return( dge )
}
```




```

# Genesets and similar

```{r}

get_genesets = function(key = NULL){
  X = read.csv(file.path(proj_dir, "tables", "genesets.csv"), stringsAsFactors = F) %>% lapply(setdiff, "")
  if(is.null(key)){
    return(X)
  }
  key %<>% make.names
  X[[key]]  %>% make.names
}



get_hox_genes = function(dge){
  grep("^hox", AvailableData(dge), ignore.case = T, value = T)
}



```





